---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# simfish

simulate fish tracks with/without acoustic detections in semi-realistic environments

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)
<!-- badges: end -->

{simfish} is an R package that facilitates simulation of fish tracks in semi-realistic environments (coastal oceans, fjords, and lakes). A user-supplied raster defining the water body - land boundaries is used to constrain simulated tracks. Fish movements are simulated (currently) as either a correlated random walk or a biased & correlated random walk, with the bias toward a defined Centre-of-Attraction. A potential function is used to ensure the tracks avoid land. The package can also be used to simulate detections of acoustically-tagged fish by acoustic receivers at user-defined locations.

## Installation

You can install simfish from [GitHub](https://github.com/ianjonsen/simfish) with:

``` r
# install.packages("remotes")
remotes::install_github("ianjonsen/simfish")
```

## Example

```{r load pkgs, message=FALSE}
require(simfish, quietly = TRUE)
require(raster, quietly = TRUE)
require(tidyverse, quietly = TRUE)
require(sf, quietly = TRUE)
require(stars, quietly = TRUE)
```

## Create an environment to simulate fish movements
```{r set up raster files for simulation, eval=FALSE}
## create raster using rnaturalearth polygon data
land <- rnaturalearth::ne_countries(scale = 10, returnclass = "sf") %>%
    sf::st_transform(crs = 4326) %>%
    sf::st_crop(xmin=-68, ymin=43.5, xmax=-52, ymax=52) %>%
    sf::st_make_valid()

## rasterise
land <- as(land, "Spatial")

## rasterize at a high resolution for a pretty map
##  a lower resolution will run faster
land <- raster::raster(crs = crs(land), 
                   vals = 1, 
                   resolution = c(0.01, 0.01), 
                   ext = extent(c(-68, -52, 43.5, 52))) %>% 
  raster::rasterize(land, .)

## reproject to Mercator grid in km
## in principle, any projection will work as long as the units are in km
ext <- raster::projectExtent(land, crs = "+proj=merc +datum=WGS84 +units=km")
land <- raster::projectRaster(land, ext)
land[land > 1] <- 1
land2 <- land # make a copy

## set water = 1 & all land values to NA
land2[is.na(land2)] <- -1
land2[land2 == 1] <- NA
land2[land2 == -1] <- 1

## calculate gradient rasters - these are needed to keep fish off land
dist <- raster::distance(land2)
grad <- ctmcmove::rast.grad(dist)
grad <- raster::stack(grad$rast.grad.x, grad$rast.grad.y)

## write files for repeated use
writeRaster(grad, file = "data/grad.grd", overwrite = TRUE)
writeRaster(land, file = "data/land.grd", overwrite = TRUE)
```

```{r get land, echo=FALSE}
land <- raster("data/land.grd")
```

## Here is the land raster we've just created
```{r map land}

raster::plot(land)
```


## Set up the simulation
```{r setup simulation}
## first create a list with the required data
x <- list(land = raster("data/land.grd"), 
          grad = stack("data/grad.grd"), 
          prj = "+proj=merc +datum=WGS84 +units=km")

## then parameterize the simulation
my.par <- sim_par(
  N = 70*24,  # number of simulation time steps (60 days x 24 h)
  start = c(-7260, 5930), # start location for simulated fish
  start.dt = ISOdatetime(2023,03,01,12,00,00, tz="UTC"), # start date-time
  time.step = 60, # time step in minutes
  coa = c(-6250, 6000), # location of centre-of-attraction for biased correlated random walk (can be NA)
  nu = 0.4, # strength of attraction to CoA (range: 0 - infinity)
  fl = 0.15, # fish fork length in m
  rho = 0.85, # angular dispersion param for move directions (range 0 - 1)
  bl = 2, # move speed (body-lengths per s)
  beta = c(-2, -2), # potential function params; must be -ve to keep off land;
  # larger -ve numbers will cause the track to jump (unrealistically) big distances when encountering land,
  # possibly crossing narrow land masses
  buffer = 0.5
)
```

## Simulate a single track as a biased & correlated random walk
```{r run simulation, eval=FALSE}
## simulate a single track
out <- sim_fish(id = 1, 
                data = x, 
                mpar = my.par)
```

```{r run simulation2, echo=FALSE}
## specify random seed to ensure same track is simulated each time this .Rmd runs
##  DO NOT do this for 'real' applications!
set.seed(300)

## simulate a single track
out <- sim_fish(id = 1, 
                data = x, 
                mpar = my.par,
                pb = FALSE) # turn off progress bar for tidy Rmd result
```

## Now simulate acoustic detections on a receiver array - we'll use the Ocean Tracking Network's Cabot Strait receiver line
```{r, message=FALSE}
tmp <- tempfile(fileext = ".csv")
download.file("https://members.oceantrack.org//geoserver/wfs?request=getfeature&service=wfs&outputFormat=CSV&typename=otn:stations_receivers&cql_filter=collectioncode+EQ+%27CBS%27",
              destfile = tmp)
recLocs <- read_csv(tmp) %>%
  filter(deploy_date >= "2020-01-01")

xy <- recLocs %>%
  st_as_sf(., coords = c("stn_long", "stn_lat"), crs = 4326) %>%
  st_transform(crs = "+proj=merc +datum=WGS84 +units=km") %>%
  st_coordinates() %>%
  as.data.frame()
recLocs <- bind_cols(recLocs, xy) %>%
  select(id = station_name, x = X, y = Y) %>%
  mutate(z = 1)

## add receiver locations to data
x$recLocs <- recLocs

## simulate detections of a V13 acoustic tag
##   parameters (intercept, slope) from a logistic regression of distance on 
##   detection rate
my.par$pdrf <- c(5, -0.01) 
out <- out %>%
  sim_detect(data = x)

```

## Visualise simulated track & detections
```{r map results}
ggplot() +
  stars::geom_stars(data = stars::st_as_stars(x$land)) +
  geom_point(data = x$recLocs,
              aes(x, y), 
              shape = 19, 
              colour = "firebrick",
              alpha = 0.6,
             size = 0.5) +
  geom_point(data = out$sim, 
             aes(x, y), 
             size = 0.1, 
             colour = "orange") +
  geom_point(data = out$detect,
             aes(recv_x, recv_y),
             shape = 19,
             size = 3,
             colour = "hotpink") +
  geom_point(data = with(out$params, data.frame(x = coa[1], y = coa[2])),
             aes(x, y),
             shape = 17,
             size = 2, 
             colour = "dodgerblue") +
  theme_minimal() +
  theme(legend.position = "none") + 
  labs(x = element_blank(),
       y = element_blank()) +
  coord_sf(expand = FALSE)

```


