---
title: "Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(simfish)
```

`simfish` can be used to simulate movement tracks and acoustic detections in featureless and semi-realistic environments. Users can generate an environmental raster - a raster of land and water features - using the built-in `generate_env` function. This function generates both a land raster and two gradient rasters that are used to keep simulated tracks off of land via a potential function. Alternatively, users may supply their own environmental rasters  (e.g., generated from an ESRI shapefile or similar) when a fine-scale simulation in a high-resolution environment is required. 

### Parameterizing the simulation
Simulations are parameterized using the `sim_par` function. `sim_par()` has default values for all parameters, which can be checked by typing `?sim_par`. 
